Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-09-27 docs/_tex/index.tex
add results

The primary objective of this study was to develop a method for the
accurate classification of macrophytes at low tide on tidal flats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral
resolution. The discrimination of seagrasses from green macroalgae
presents is spectrally challenging \citetext{\citealp[
]{oiry2021using}; \citealp[
]{bannari2022}; \citealp{veettil2020opportunities}}. These two
macrophytes share a similar pigment composition: chlorophyll-a (common
to all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Their spectreal responses could be
close, particularly at a multispectral resolution. Seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone. In this study the risk of spectral
confusion was avoided with a machine-learning approach exploiting a
neural networks classifier. Our drone flights and a recent study based
on \emph{in situ} radiometry, suggested that a sensor with at least
eight spectral bands ranging from 500 to 850 nm and including a green
band at 530 nm and a RedEdge band at 730 nm was crucial to accurately
discriminate green macroalgae from seagrasses \citep{Davies2023}.

@@

The ability to differentiate between various types of vegetation plays a
critical role in ecological monitoring and coastal management
\citep{WFD2000}. By distinguishing between seagrasses and macroalgae,
our approach facilitates targeted conservation strategies, enabling more
effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120
m), a nearly one-to-one relationship was observed, with a Root Mean
Square Error (RMSE) of 0.02 (Figure~\ref{fig-CompareRef}). This result
indicates that the reflectance measured by remote sensing (RS) sensors
is not significantly influenced by pixel size. This finding is valuable
for integrating drone-based data into larger-scale mapping projects
(e.g., combining satellite and drone mapping in side-by-side analyses).
The consistency of reflectance across altitudes suggests that drones can
be effectively used for finer-scale mapping without compromising data
accuracy when merging with other platforms. However, it was observed
that there is an underestimation of the infrared part of the spectra in
the high-altitude dataset (Figure~\ref{fig-CompareRef}). Such disparity
in infrared reflectance may stem from temporal differences between the
flights, possibly resulting in a slightly drier intertidal area and
consequently higher infrared reflectance. This disparity poses an issue
for the methodology followed in the present study, relying solely on one
flight height for training. To address this issue, we employed min/max
standardized reflectance spectra as predictors for the model
Equation~\ref{eq-std}. This approach allowed us to eliminate the slight
reflectance difference between the flights (Figure~\ref{fig-CompareRef}
B) and to focus on the shape of the spectra in the visible part of the
electromagnetic spectra, where different pigmentation are associated to
taxonomic dignostic features. In contrast to subtidal seagrasses, which
maintain a relatively constant biomass throughout the year, intertidal
seagrasses, like the one studied in this work, exhibit strong seasonal
phenology \citep{davies2024sentinel}. At some sites, they completely
disappear during the winter and reach their peak above-ground biomass in
the summer and early fall. Along with these seasonal changes in biomass,
the pigment composition and ratios also vary throughout the year,
reflecting the plants' adaptations to different environmental conditions
\citetext{\citealp[ ]{bargain2013seasonal}; \citealp{legare2022remote}}.
Standardization of spectral signatures helps to mitigate the impact of
changing biomass on the spectral profile, enabling the development of a
model that can reliably predict vegetation across different geographical
locations and seasons. This approach allows for consistent
classification of vegetation despite variations in biomass and
fluctuations in light conditions, providing a robust tool for monitoring
and predicting vegetation dynamics \citetext{\citealp[
]{fyfe2003spatial}; \citealp[
]{COSTA2021107018}; \citealp{piaser2023impact}}. We found 90 \% seagrass
cover was necessary at an 8 cm resolution for confident prediction of
its presence (Figure~\ref{fig-upscaling}). However, due to the strong
phenology of intertidal seagrass meadows in Europe, the period when a
meadow is well-established can be temporally restricted, limiting the
ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive
species development, coastal erosion, and sealevel rise are expected to
continue impacting coastal ecosystems in the future
\citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}}. and the demand for
meaningful and efficient monitoring of coastal habitats has never been
higher\citetext{\citealp[ ]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of intertidal seagrass and
green macroalgae from other intertidal vegetation classes, highlight the
potential of drone-based remote sensing to support diverse applications,
from conservation of biodiversity to climate change adaptation
strategies.

Because of coastal eutrophication, macroalgal blooms are becoming
increasingly common in many regions around the world \citetext{\citealp[
]{sutton2011european}; \citealp{ye2011green}}. These blooms can have
negative impacts on human health and local economic activities,
including human health, fishing and aquaculture, tourism, and
recreational activities \citetext{\citealp[
]{villares1999nitrogen}; \citealp{ye2011green}}. The first green tide
events (i.e.~bloom of green macroalgae of the genus \emph{Ulva}) were
reported in Brittany, France, back in the 1970s and have since been a
concern for local stakeholders and economic activities
\citep{menesguen2018marees}. Some regions of the world have witnessed an
increase in brown macroalgae blooms, predominantly involving algae of
the genus \emph{Sargassum} washing along the Caribbean coastlines
\citep{louime2017sargassum}, and more recently \emph{Rugulopteryx
okamurea} in southern Europe \citep{Roca2022}. Satellite remote sensing
has proven to be a valuable tool for mapping the spatial and temporal
extent of macroalgal blooms worldwide. However, due to limitations in
spatial resolution, it can only effectively map well-developed blooms
\citetext{\citealp[ ]{rs13081408}; \citealp[
]{klemas2012remote}; \citealp{haro2023biointertidal}}. High spatial
resolution drone imagery, coupled with an accurate classification
algorithm, could be used to map the early stages of macroalgal blooms in
areas known to have regular blooms or in new sites. Indeed, this
approach could provide early warning alerts to local managers.

Employing traditional sampling methods to monitor coastal ecosystems is
time and resource-intensive, and the findings are often difficult to
scale-up. Earth Observation can bridge this gap and meet the needs for
systematic monitoring coastal ecosystems over large areas
\citep{papathanasopoulou2019satellite}. The retrieval of Essential
Biodiversity Variables and Essential Ocean Variables through satellite
observations has been increasingly common, enabling comprehensive
monitoring of entire ecosystems over extended time periods
\citetext{\citealp[ ]{ratnarajah2023monitoring}; \citealp{Zoffoli2021}}.
The Water Framework Directive \citep{WFD2000} mandates the achievement
and maintenance of ``good ecological status'' for all European waters,
which necessitates a comprehensive understanding and monitoring of
aquatic ecosystems, including coastal habitats like seagrass beds
\citetext{\citealp[ ]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.

Effective and efficient monitoring tools are essential for identifying
the impacts of human activities and natural changes on coastal
ecosystems. On-demand, multispectral drone observations at very high
spatial-resolution provide a novel and powerful tool to rapidly and
accurately acquire ground truth satellite data. Spatially resolved data
are indeed critical for calibrating and validating satellite remote
sensing observations, thereby enhancing our capacity to monitor vast
coastal areas. A perspective of this work could be to develop a similar
classifier for satellite images (e.g.~Sentinel-2), and test whether the
discrimination between seagrass and green macroalgae is still
operational at a coarser spatial resolution. The integration of drone
technology facilitates a scalable approach to environmental
surveillance, offering significant advancements in the spatial and
temporal resolution of data collection. This, in turn, supports the EU
WFD's objectives by enabling more informed and timely management
decisions for the conservation and restoration of aquatic ecosystems.

>>>>

The primary objective of this study was to develop a method for the
accurate classification of macrophytes at low tide on tidal flats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral
resolution. The discrimination of seagrasses from green macroalgae
presents is spectrally challenging \citetext{\citealp[
]{oiry2021using}; \citealp[
]{bannari2022}; \citealp{veettil2020opportunities}}. These two
macrophytes share a similar pigment composition: chlorophyll-a (common
to all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Their spectral responses could be
close, particularly at a multispectral resolution. Seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone. In this study the risk of spectral
confusion was avoided with a machine-learning approach exploiting a
neural networks classifier. Our drone flights and a recent study based
on \emph{in situ} radiometry, suggested that a sensor with at least
eight spectral bands ranging from 500 to 850 nm and including a green
band at 530 nm and a RedEdge band at 730 nm was crucial to accurately
discriminate green macroalgae from seagrasses \citep{Davies2023}.

@@

The ability to differentiate between various types of vegetation plays a
critical role in ecological monitoring and coastal management
\citep{WFD2000}. By distinguishing between seagrasses and macroalgae,
our approach facilitates targeted conservation strategies, enabling more
effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120
m), a nearly one-to-one relationship was observed, with a Root Mean
Square Error (RMSE) of 0.02 (Figure~\ref{fig-CompareRef}). This result
indicates that the reflectance measured by remote sensing (RS) sensors
is not significantly influenced by pixel size. This finding is valuable
for integrating drone-based data into larger-scale mapping projects
(e.g., combining satellite and drone mapping in side-by-side analyses).
The consistency of reflectance across altitudes suggests that drones can
be effectively used for finer-scale mapping without compromising data
accuracy when merging with other platforms. However, it was observed
that there is an underestimation of the infrared part of the spectra in
the high-altitude dataset (Figure~\ref{fig-CompareRef}). Such disparity
in infrared reflectance may stem from temporal differences between the
flights, possibly resulting in a slightly drier intertidal area and
consequently higher infrared reflectance. This disparity poses an issue
for the methodology followed in the present study, relying solely on one
flight height for training. To address this issue, we employed min/max
standardized reflectance spectra as predictors for the model
Equation~\ref{eq-std}. This approach allowed us to eliminate the slight
reflectance difference between the flights (Figure~\ref{fig-CompareRef}
B) and to focus on the shape of the spectra in the visible part of the
electromagnetic spectra, where different pigmentation are associated to
taxonomic diagnostic features. In contrast to subtidal seagrasses, which
maintain a relatively constant biomass throughout the year, intertidal
seagrasses, like the one studied in this work, exhibit strong seasonal
phenology \citep{davies2024sentinel}. At some sites, they completely
disappear during the winter and reach their peak above-ground biomass in
the summer and early fall. Along with these seasonal changes in biomass,
the pigment composition and ratios also vary throughout the year,
reflecting the plants' adaptations to different environmental conditions
\citetext{\citealp[ ]{bargain2013seasonal}; \citealp{legare2022remote}}.
Standardization of spectral signatures helps to mitigate the impact of
changing biomass on the spectral profile, enabling the development of a
model that can reliably predict vegetation across different geographical
locations and seasons. This approach allows for consistent
classification of vegetation despite variations in biomass and
fluctuations in light conditions, providing a robust tool for monitoring
and predicting vegetation dynamics \citetext{\citealp[
]{fyfe2003spatial}; \citealp[
]{COSTA2021107018}; \citealp{piaser2023impact}}. We found 90 \% seagrass
cover was necessary at an 8 cm resolution for confident prediction of
its presence (Figure~\ref{fig-upscaling}). However, due to the strong
phenology of intertidal seagrass meadows in Europe, the period when a
meadow is well-established can be temporally restricted, limiting the
ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive
species development, coastal erosion, and sea level rise are expected to
continue impacting coastal ecosystems in the future
\citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}} and the demand for
meaningful and efficient monitoring of coastal habitats has never been
higher\citetext{\citealp[ ]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of intertidal seagrass and
green macroalgae from other intertidal vegetation classes, highlight the
potential of drone-based remote sensing to support diverse applications,
from conservation of biodiversity to climate change adaptation
strategies.

Because of coastal eutrophication, macroalgal blooms are becoming
increasingly common in many regions around the world \citetext{\citealp[
]{sutton2011european}; \citealp{ye2011green}}. These blooms can have
negative impacts on human health and local economic activities,
including human health, fishing and aquaculture, tourism, and
recreational activities \citetext{\citealp[
]{villares1999nitrogen}; \citealp{ye2011green}}. The first green tide
events (i.e.~bloom of green macroalgae of the genus \emph{Ulva}) were
reported in Brittany, France, back in the 1970s and have since been a
concern for local stakeholders and economic activities
\citep{menesguen2018marees}. Some regions of the world have witnessed an
increase in brown macroalgae blooms, predominantly involving algae of
the genus \emph{Sargassum} washing along the Caribbean coastlines
\citep{louime2017sargassum}, and more recently \emph{Rugulopteryx
okamurea} in southern Europe \citep{Roca2022}. Satellite remote sensing
has proven to be a valuable tool for mapping the spatial and temporal
extent of macroalgal blooms worldwide. However, due to limitations in
spatial resolution, it can only effectively map well-developed blooms
\citetext{\citealp[ ]{rs13081408}; \citealp[
]{klemas2012remote}; \citealp{haro2023biointertidal}}. High spatial
resolution drone imagery, coupled with an accurate classification
algorithm, could be used to map the early stages of macroalgal blooms in
areas known to have regular blooms or in new sites. Indeed, this
approach could provide early warning alerts to local managers and
complimentary to traditional sampling methods to monitor coastal
ecosystems. These methods are generally time and resource-intensive, and
the findings are often difficult to scale-up when applied alone. Earth
Observation can bridge this gap and meet the needs for systematic
monitoring coastal ecosystems over large areas
\citep{papathanasopoulou2019satellite}. The retrieval of Essential
Biodiversity Variables and Essential Ocean Variables through satellite
observations has been increasingly common, enabling comprehensive
monitoring of entire ecosystems over extended time periods
\citetext{\citealp[ ]{ratnarajah2023monitoring}; \citealp{Zoffoli2021}}.
The Water Framework Directive \citep{WFD2000} mandates the achievement
and maintenance of ``good ecological status'' for all European waters,
which necessitates a comprehensive understanding and monitoring of
aquatic ecosystems, including coastal habitats like seagrass beds
\citetext{\citealp[ ]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.

Effective and efficient monitoring tools are essential for identifying
the impacts of human activities and natural changes on coastal
ecosystems. On-demand, multispectral drone observations at very high
spatial-resolution provide a novel and powerful tool to rapidly and
accurately acquire ground truth data which can be used to develop
machine-learning algorithm for satellite sensors
\citep{davies2024intertidal}. Spatially resolved data are indeed
critical for calibrating and validating satellite remote sensing
observations, thereby enhancing our capacity to monitor vast coastal
areas. A perspective of this work could be to develop a similar
classifier for satellite images (e.g.~Sentinel-2), and test whether the
discrimination between seagrass and green macroalgae is still
operational at a coarser spatial resolution. The integration of drone
technology facilitates a scalable approach to environmental surveillance
while taking into account patchiness of secondary vegetation, offering
significant advancements in the spatial and temporal resolution of data
collection. This, in turn, supports the EU WFD's objectives by enabling
more informed and timely management decisions for the conservation and
restoration of aquatic ecosystems.

### 2024-09-27 index.qmd
add results

The primary objective of this study was to develop a method for the accurate classification of macrophytes at low tide on tidal flats, specifically focusing on distinguishing between Chlorophyceae (green macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral resolution. The discrimination of seagrasses from green macroalgae presents is spectrally challenging [@oiry2021using ; @bannari2022 ; @veettil2020opportunities]. These two macrophytes share a similar pigment composition: chlorophyll-a (common to all vegetation types), chlorophyll-b (an additional photosynthetic pigment), and accessory carotenoids such as zeaxanthin, lutein and neoxanthin (@fig-Pigm). Their spectreal responses could be close, particularly at a multispectral resolution. Seagrass and green macroalgae frequently co-occur in intertidal areas, and can intermingle within a remote sensing pixel if the spatial resolution is too low. Here, the issue of intra-pixel mixing was resolved thanks to the very high spatial resolution of the drone. In this study the risk of spectral confusion was avoided with a machine-learning approach exploiting a neural networks classifier. Our drone flights and a recent study based on *in situ* radiometry, suggested that a sensor with at least eight spectral bands ranging from 500 to 850 nm and including a green band at 530 nm and a RedEdge band at 730 nm was crucial to accurately discriminate green macroalgae from seagrasses [@Davies2023].

@@

The ability to differentiate between various types of vegetation plays a critical role in ecological monitoring and coastal management [@WFD2000]. By distinguishing between seagrasses and macroalgae, our approach facilitates targeted conservation strategies, enabling more effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120 m), a nearly one-to-one relationship was observed, with a Root Mean Square Error (RMSE) of 0.02 (@fig-CompareRef). This result indicates that the reflectance measured by remote sensing (RS) sensors is not significantly influenced by pixel size. This finding is valuable for integrating drone-based data into larger-scale mapping projects (e.g., combining satellite and drone mapping in side-by-side analyses). The consistency of reflectance across altitudes suggests that drones can be effectively used for finer-scale mapping without compromising data accuracy when merging with other platforms. However, it was observed that there is an underestimation of the infrared part of the spectra in the high-altitude dataset (@fig-CompareRef). Such disparity in infrared reflectance may stem from temporal differences between the flights, possibly resulting in a slightly drier intertidal area and consequently higher infrared reflectance. This disparity poses an issue for the methodology followed in the present study, relying solely on one flight height for training. To address this issue, we employed min/max standardized reflectance spectra as predictors for the model [@eq-std ; @Cao2017]. This approach allowed us to eliminate the slight reflectance difference between the flights (@fig-CompareRef B) and to focus on the shape of the spectra in the visible part of the electromagnetic spectra, where different pigmentation are associated to taxonomic dignostic features. In contrast to subtidal seagrasses, which maintain a relatively constant biomass throughout the year, intertidal seagrasses, like the one studied in this work, exhibit strong seasonal phenology [@davies2024sentinel]. At some sites, they completely disappear during the winter and reach their peak above-ground biomass in the summer and early fall. Along with these seasonal changes in biomass, the pigment composition and ratios also vary throughout the year, reflecting the plants' adaptations to different environmental conditions [@bargain2013seasonal ; @legare2022remote]. Standardization of spectral signatures helps to mitigate the impact of changing biomass on the spectral profile, enabling the development of a model that can reliably predict vegetation across different geographical locations and seasons. This approach allows for consistent classification of vegetation despite variations in biomass and fluctuations in light conditions, providing a robust tool for monitoring and predicting vegetation dynamics [@fyfe2003spatial ; @COSTA2021107018 ; @piaser2023impact]. We found `r round((lm_magno$coefficients[[2]]*85+lm_magno$coefficients[[1]])*100,0)` % seagrass cover was necessary at an 8 cm resolution for confident prediction of its presence (@fig-upscaling). However, due to the strong phenology of intertidal seagrass meadows in Europe, the period when a meadow is well-established can be temporally restricted, limiting the ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive species development, coastal erosion, and sealevel rise are expected to continue impacting coastal ecosystems in the future [@SCHIBALSKI2022101414; @holon2018predictive ; @marquet2024global]. and the demand for meaningful and efficient monitoring of coastal habitats has never been higher[@muller2018satellite ; @villalobos2023remote ; @oiry2021using]. Our findings, particularly the improved discrimination of intertidal seagrass and green macroalgae from other intertidal vegetation classes, highlight the potential of drone-based remote sensing to support diverse applications, from conservation of biodiversity to climate change adaptation strategies.

Because of coastal eutrophication, macroalgal blooms are becoming increasingly common in many regions around the world [@sutton2011european ; @ye2011green]. These blooms can have negative impacts on human health and local economic activities, including human health, fishing and aquaculture, tourism, and recreational activities [@villares1999nitrogen ; @ye2011green]. The first green tide events (i.e. bloom of green macroalgae of the genus *Ulva*) were reported in Brittany, France, back in the 1970s and have since been a concern for local stakeholders and economic activities [@menesguen2018marees]. Some regions of the world have witnessed an increase in brown macroalgae blooms, predominantly involving algae of the genus *Sargassum* washing along the Caribbean coastlines [@louime2017sargassum], and more recently *Rugulopteryx okamurea* in southern Europe [@Roca2022]. Satellite remote sensing has proven to be a valuable tool for mapping the spatial and temporal extent of macroalgal blooms worldwide. However, due to limitations in spatial resolution, it can only effectively map well-developed blooms [@rs13081408 ; @klemas2012remote ; @haro2023biointertidal]. High spatial resolution drone imagery, coupled with an accurate classification algorithm, could be used to map the early stages of macroalgal blooms in areas known to have regular blooms or in new sites. Indeed, this approach could provide early warning alerts to local managers.

Employing traditional sampling methods to monitor coastal ecosystems is time and resource-intensive, and the findings are often difficult to scale-up. Earth Observation can bridge this gap and meet the needs for systematic monitoring coastal ecosystems over large areas [@papathanasopoulou2019satellite]. The retrieval of Essential Biodiversity Variables and Essential Ocean Variables through satellite observations has been increasingly common, enabling comprehensive monitoring of entire ecosystems over extended time periods [@ratnarajah2023monitoring ; @Zoffoli2021]. The Water Framework Directive [@WFD2000] mandates the achievement and maintenance of "good ecological status" for all European waters, which necessitates a comprehensive understanding and monitoring of aquatic ecosystems, including coastal habitats like seagrass beds [@foden2007angiosperms ; @nordlund2024one ; @Zoffoli2021].

Effective and efficient monitoring tools are essential for identifying the impacts of human activities and natural changes on coastal ecosystems. On-demand, multispectral drone observations at very high spatial-resolution provide a novel and powerful tool to rapidly and accurately acquire ground truth satellite data. Spatially resolved data are indeed critical for calibrating and validating satellite remote sensing observations, thereby enhancing our capacity to monitor vast coastal areas. A perspective of this work could be to develop a similar classifier for satellite images (e.g. Sentinel-2), and test whether the discrimination between seagrass and green macroalgae is still operational at a coarser spatial resolution. The integration of drone technology facilitates a scalable approach to environmental surveillance, offering significant advancements in the spatial and temporal resolution of data collection. This, in turn, supports the EU WFD's objectives by enabling more informed and timely management decisions for the conservation and restoration of aquatic ecosystems.

>>>>

The primary objective of this study was to develop a method for the accurate classification of macrophytes at low tide on tidal flats, specifically focusing on distinguishing between Chlorophyceae (green macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral resolution. The discrimination of seagrasses from green macroalgae presents is spectrally challenging [@oiry2021using ; @bannari2022 ; @veettil2020opportunities]. These two macrophytes share a similar pigment composition: chlorophyll-a (common to all vegetation types), chlorophyll-b (an additional photosynthetic pigment), and accessory carotenoids such as zeaxanthin, lutein and neoxanthin (@fig-Pigm). Their spectral responses could be close, particularly at a multispectral resolution. Seagrass and green macroalgae frequently co-occur in intertidal areas, and can intermingle within a remote sensing pixel if the spatial resolution is too low. Here, the issue of intra-pixel mixing was resolved thanks to the very high spatial resolution of the drone. In this study the risk of spectral confusion was avoided with a machine-learning approach exploiting a neural networks classifier. Our drone flights and a recent study based on *in situ* radiometry, suggested that a sensor with at least eight spectral bands ranging from 500 to 850 nm and including a green band at 530 nm and a RedEdge band at 730 nm was crucial to accurately discriminate green macroalgae from seagrasses [@Davies2023].

@@

The ability to differentiate between various types of vegetation plays a critical role in ecological monitoring and coastal management [@WFD2000]. By distinguishing between seagrasses and macroalgae, our approach facilitates targeted conservation strategies, enabling more effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120 m), a nearly one-to-one relationship was observed, with a Root Mean Square Error (RMSE) of 0.02 (@fig-CompareRef). This result indicates that the reflectance measured by remote sensing (RS) sensors is not significantly influenced by pixel size. This finding is valuable for integrating drone-based data into larger-scale mapping projects (e.g., combining satellite and drone mapping in side-by-side analyses). The consistency of reflectance across altitudes suggests that drones can be effectively used for finer-scale mapping without compromising data accuracy when merging with other platforms. However, it was observed that there is an underestimation of the infrared part of the spectra in the high-altitude dataset (@fig-CompareRef). Such disparity in infrared reflectance may stem from temporal differences between the flights, possibly resulting in a slightly drier intertidal area and consequently higher infrared reflectance. This disparity poses an issue for the methodology followed in the present study, relying solely on one flight height for training. To address this issue, we employed min/max standardized reflectance spectra as predictors for the model [@eq-std ; @Cao2017]. This approach allowed us to eliminate the slight reflectance difference between the flights (@fig-CompareRef B) and to focus on the shape of the spectra in the visible part of the electromagnetic spectra, where different pigmentation are associated to taxonomic diagnostic features. In contrast to subtidal seagrasses, which maintain a relatively constant biomass throughout the year, intertidal seagrasses, like the one studied in this work, exhibit strong seasonal phenology [@davies2024sentinel]. At some sites, they completely disappear during the winter and reach their peak above-ground biomass in the summer and early fall. Along with these seasonal changes in biomass, the pigment composition and ratios also vary throughout the year, reflecting the plants' adaptations to different environmental conditions [@bargain2013seasonal ; @legare2022remote]. Standardization of spectral signatures helps to mitigate the impact of changing biomass on the spectral profile, enabling the development of a model that can reliably predict vegetation across different geographical locations and seasons. This approach allows for consistent classification of vegetation despite variations in biomass and fluctuations in light conditions, providing a robust tool for monitoring and predicting vegetation dynamics [@fyfe2003spatial ; @COSTA2021107018 ; @piaser2023impact]. We found `r round((lm_magno$coefficients[[2]]*85+lm_magno$coefficients[[1]])*100,0)` % seagrass cover was necessary at an 8 cm resolution for confident prediction of its presence (@fig-upscaling). However, due to the strong phenology of intertidal seagrass meadows in Europe, the period when a meadow is well-established can be temporally restricted, limiting the ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive species development, coastal erosion, and sea level rise are expected to continue impacting coastal ecosystems in the future [@SCHIBALSKI2022101414; @holon2018predictive ; @marquet2024global] and the demand for meaningful and efficient monitoring of coastal habitats has never been higher[@muller2018satellite ; @villalobos2023remote ; @oiry2021using]. Our findings, particularly the improved discrimination of intertidal seagrass and green macroalgae from other intertidal vegetation classes, highlight the potential of drone-based remote sensing to support diverse applications, from conservation of biodiversity to climate change adaptation strategies.

Because of coastal eutrophication, macroalgal blooms are becoming increasingly common in many regions around the world [@sutton2011european ; @ye2011green]. These blooms can have negative impacts on human health and local economic activities, including human health, fishing and aquaculture, tourism, and recreational activities [@villares1999nitrogen ; @ye2011green]. The first green tide events (i.e. bloom of green macroalgae of the genus *Ulva*) were reported in Brittany, France, back in the 1970s and have since been a concern for local stakeholders and economic activities [@menesguen2018marees]. Some regions of the world have witnessed an increase in brown macroalgae blooms, predominantly involving algae of the genus *Sargassum* washing along the Caribbean coastlines [@louime2017sargassum], and more recently *Rugulopteryx okamurea* in southern Europe [@Roca2022]. Satellite remote sensing has proven to be a valuable tool for mapping the spatial and temporal extent of macroalgal blooms worldwide. However, due to limitations in spatial resolution, it can only effectively map well-developed blooms [@rs13081408 ; @klemas2012remote ; @haro2023biointertidal]. High spatial resolution drone imagery, coupled with an accurate classification algorithm, could be used to map the early stages of macroalgal blooms in areas known to have regular blooms or in new sites. Indeed, this approach could provide early warning alerts to local managers and complimentary to traditional sampling methods to monitor coastal ecosystems. These methods are generally time and resource-intensive, and the findings are often difficult to scale-up when applied alone. Earth Observation can bridge this gap and meet the needs for systematic monitoring coastal ecosystems over large areas [@papathanasopoulou2019satellite]. The retrieval of Essential Biodiversity Variables and Essential Ocean Variables through satellite observations has been increasingly common, enabling comprehensive monitoring of entire ecosystems over extended time periods [@ratnarajah2023monitoring ; @Zoffoli2021]. The Water Framework Directive [@WFD2000] mandates the achievement and maintenance of "good ecological status" for all European waters, which necessitates a comprehensive understanding and monitoring of aquatic ecosystems, including coastal habitats like seagrass beds [@foden2007angiosperms ; @nordlund2024one ; @Zoffoli2021].

Effective and efficient monitoring tools are essential for identifying the impacts of human activities and natural changes on coastal ecosystems. On-demand, multispectral drone observations at very high spatial-resolution provide a novel and powerful tool to rapidly and accurately acquire ground truth data which can be used to develop machine-learning algorithm for satellite sensors [@davies2024intertidal]. Spatially resolved data are indeed critical for calibrating and validating satellite remote sensing observations, thereby enhancing our capacity to monitor vast coastal areas. A perspective of this work could be to develop a similar classifier for satellite images (e.g. Sentinel-2), and test whether the discrimination between seagrass and green macroalgae is still operational at a coarser spatial resolution. The integration of drone technology facilitates a scalable approach to environmental surveillance while taking into account patchiness of secondary vegetation, offering significant advancements in the spatial and temporal resolution of data collection. This, in turn, supports the EU WFD's objectives by enabling more informed and timely management decisions for the conservation and restoration of aquatic ecosystems.
